---
title: "BaSiC-simulation"
author: "Wei Sun"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This R markdown file modify the basic simulation anlaysis in ex_simulation.Rmd to perform a sensitivity analyses: adjusting the read-depth cutoff to call mutations to be 16.

```{r libraries, warning = FALSE, message = FALSE}
library(basic)
library(ggcorrplot)
library(gridExtra)
```

### Analysis without copy number alterations

We first evaluate the simulated data without any copy number alterations. 

```{r analysis}
load("basic/data/simulation_SCS_bulk.RData")

b.trec    = simu.bulk$trec
b.asrec   = simu.bulk$asrec
b.cn      = rep(2, length(b.trec))

mutations = simu.SCS$observed.mutation.nomissing
trec      = simu.SCS$trec
asrec     = simu.SCS$asrec

dim(mutations)
mutations[1:2,1:3]

mutations[which(trec <= 16)] = NA
table(mutations, useNA="ifany")

cBSC = cluster.BSC(b.cn, b.asrec, b.trec, mutations, asrec, trec, nclusters=2:5)

Ds = simu.SCS$observed.mutation
Es = simu.SCS$true.mutations
cBSC$summary.cluster
```

From the summary of above analysis, both AIC and BIC suggest that there are four clusters. Next we visualize the distribution of estimated mutation status (```Ehats```) vs. the true mutation status (```E```). 

```{r plot1, fig.width=6, fig.height=6}
par(mfrow=c(2,2), mar=c(4,4,2,1))
hist(c(Ds)[which(c(Es)==0)], main="Ds | Es=0", xlab="", breaks=20)
hist(c(Ds)[which(c(Es)==1)], main="Ds | Es=1", xlab="", breaks=20)
hist(c(cBSC$n3$E)[which(c(Es)==0)], main="Ehats | Es=0", xlab="", breaks=20)
hist(c(cBSC$n3$E)[which(c(Es)==1)], main="Ehats | Es=1", xlab="", breaks=20)
```

Visualize the whole mutation matrix for observed mutations, true mutations and estimated mutations. 
```{r plot2, fig.width=9, fig.height=4.5}
par(mfrow=c(1,3), mar=c(4,4,4,1))
image(t(Ds), col=c("grey", "darkred"), axes = FALSE, xlab="Cells", 
      ylab="Mutations", main="observed mutations")
image(t(Es), col=c("grey", "darkred"), axes = FALSE, xlab="Cells", 
      ylab="Mutations", main="true mutations")
image(t(cBSC$n3$E), col=c("grey", "darkred"), axes = FALSE, xlab="Cells", 
      ylab="Mutations", main="esimated mutations")
```

### Analysis with copy number alterations
Next we conduct similar analysis using the simulated data with copy number alterations.
```{r load_data_cna}
load("basic/data/simulation_SCS_bulk_with_cn.RData")

b.trec    = simu.bulk$trec
b.asrec   = simu.bulk$asrec
b.cn      = simu.bulk$cn

trec      = simu.SCS$trec
asrec     = simu.SCS$asrec
mutations = simu.SCS$observed.mutation.nomissing

dim(mutations)
mutations[1:2,1:3]

mutations[which(trec <= 16)] = NA
table(c(mutations), useNA="ifany")
```

Single cell only analysis.
```{r analysis_cna_sc}
cd1 = cluster.SCS(mutations, asrec, trec, nclusters=2:5)
cd1$summary.cluster
```

Joint analysis of bulk and scDNA-seq with CNA information.

```{r analysis_cna_basic}
cBSC  = cluster.BSC(b.cn, b.asrec, b.trec, mutations, asrec, trec, 
                    nclusters=2:5)
cBSC$summary.cluster

b.cn  = rep(2, length(b.asrec))
cBSC2 = cluster.BSC(b.cn, b.asrec, b.trec, mutations, asrec, trec,
                    nclusters=2:5)
cBSC2$summary.cluster
```

We compare the cluster mean estimations (```gamma```'s) versus the its true values. 

```{r compare_cluster_mean, fig.width=6, fig.height=8}
gamma.pi = 0.95
M0 = 20

scGroups = rep(NA, M0)
scGroups[1:2] = 1
scGroups[3:4] = 2
scGroups[5:12] = 3
scGroups[13:20] = 4

table(scGroups)

gamma = matrix(1-gamma.pi, nrow=4, ncol=M0)
gamma[1,] = rep(gamma.pi, M0)
gamma[2,which(scGroups >= 2)] = gamma.pi
gamma[3,which(scGroups == 3)] = gamma.pi
gamma[4,which(scGroups == 4)] = gamma.pi
gamma

g1 = ggcorrplot(t(gamma), tl.cex=10, lab = FALSE, legend.title="") + 
  ggtitle('True configuration' ) + theme(plot.title = element_text(size = 10))

g2 = ggcorrplot(cd1[["n4"]]$gamma, tl.cex=10, lab=FALSE, legend.title="") + 
  ggtitle('SCS only' ) + theme(plot.title = element_text(size = 10))

gamma2 = cBSC2[["n4"]]$gamma
g3 = ggcorrplot(gamma2, tl.cex=10, lab=FALSE, legend.title="") + 
  ggtitle('Bulk tumor + SCS, without CNA information' ) + 
  theme(plot.title = element_text(size = 10))

g4 = ggcorrplot(cBSC[["n4"]]$gamma, tl.cex=10, lab=FALSE, legend.title="") + 
  ggtitle('Bulk tumor + SCS, with CNA information' ) +
  theme(plot.title = element_text(size = 10))

grid.arrange(g1, g2, g3, g4, nrow = 4)
```

Finally we compare the cluster membership of each method versus the true cluster membership. 
```{r compare_cluster_membership, fig.width=6, fig.height=8}
round(cBSC2$n4$gammas,2)

mut1 = c(rep(1,600), rep(2,200), rep(3,100), rep(4,100))
mut2 = apply(cd1$n3$postPs, 1, which.max)
mut3 = apply(cBSC2$n4$postPs, 1, which.max)
mut4 = apply(cBSC$n4$postPs, 1, which.max)

t2 = table(mut1, mut2)
t3 = table(mut1, mut3)
t4 = table(mut1, mut4)
ts = cbind(t2, t3, t4)
ts

sum(apply(t2, 2, max))/length(mut1)
sum(apply(t3, 2, max))/length(mut1)
sum(apply(t4, 2, max))/length(mut1)
```


```{r}
sessionInfo()
```
