---
title: "BaSiC-simulation"
author: "Wei Sun"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This R markdown file modify the basic simulation anlaysis in ex_simulation.Rmd to perform a sensitivity analyses: add noise to copy number calls

```{r libraries, warning = FALSE, message = FALSE}
library(basic)
library(ggcorrplot)
library(gridExtra)
```


### Analysis with copy number alterations
We first load the simulated data with copy number alterations, and then add noise into the copy number data. Here we randomly take 10% of loci and randomly assign their copy numbers. 
```{r load_data_cna}
load("basic/data/simulation_SCS_bulk_with_cn.RData")

b.trec    = simu.bulk$trec
b.asrec   = simu.bulk$asrec
b.cn      = simu.bulk$cn

table(b.cn)
set.seed(999)
w2pertrub = sample(length(b.cn), size = round(0.1*length(b.cn)))
b.cn[w2pertrub] = sample(1:3, size=length(w2pertrub), replace=TRUE)
table(b.cn, simu.bulk$cn)

trec      = simu.SCS$trec
asrec     = simu.SCS$asrec
mutations = simu.SCS$observed.mutation

dim(mutations)
mutations[1:2,1:3]
```

Joint analysis of bulk and scDNA-seq with CNA information. Here we use suppressWarnings since when the copy number is mistakenly specified, it is possible to be incompatible to read count, which leads to a warning message. 

```{r analysis_cna_basic}
suppressWarnings({cBSC  = cluster.BSC(b.cn, b.asrec, b.trec, mutations, 
                                     asrec, trec, nclusters=2:5)})

cBSC$summary.cluster
```

We compare the cluster mean estimations (```gamma```'s) versus the its true values. 

```{r compare_cluster_mean, fig.width=6, fig.height=8}
gamma.pi = 0.95
M0 = 20

scGroups = rep(NA, M0)
scGroups[1:2] = 1
scGroups[3:4] = 2
scGroups[5:12] = 3
scGroups[13:20] = 4

table(scGroups)

gamma = matrix(1-gamma.pi, nrow=4, ncol=M0)
gamma[1,] = rep(gamma.pi, M0)
gamma[2,which(scGroups >= 2)] = gamma.pi
gamma[3,which(scGroups == 3)] = gamma.pi
gamma[4,which(scGroups == 4)] = gamma.pi
gamma

g1 = ggcorrplot(t(gamma), tl.cex=10, lab = FALSE, legend.title="") + 
  ggtitle('True configuration' ) + theme(plot.title = element_text(size = 10))

g4 = ggcorrplot(cBSC[["n4"]]$gamma, tl.cex=10, lab=FALSE, legend.title="") + 
  ggtitle('Bulk tumor + SCS, with CNA information' ) +
  theme(plot.title = element_text(size = 10))

grid.arrange(g1, g4, nrow = 2)
```

Finally we compare the cluster membership of each method versus the true cluster membership. 
```{r compare_cluster_membership, fig.width=6, fig.height=8}
mut1 = c(rep(1,600), rep(2,200), rep(3,100), rep(4,100))
mut4 = apply(cBSC$n4$postPs, 1, which.max)

t4 = table(mut1, mut4)
t4

sum(apply(t4, 2, max))/length(mut1)
```


```{r}
sessionInfo()
```
